---
title: "ND-X User Options Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ND-X User Options Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette provides detailed guidance on configuring the ND-X pipeline through its user options system. The ND-X pipeline is highly customizable, with numerous options to control each aspect of the denoising process. This guide will help you understand the available options and how to use them effectively.

## Basic Usage

The ND-X workflow functions, most notably `NDX_Process_Subject()`, accept a `user_options` parameter that controls all aspects of the denoising process. This parameter is a list with both top-level options and nested option lists for specific modules.

```{r eval=FALSE}
library(fmridenoise)

# Example of setting user options
user_options <- list(
  # Top-level options
  max_passes = 3,
  min_des_gain_convergence = 0.005,
  task_regressor_names_for_extraction = c("face", "house"),
  
  # Module-specific options
  opts_pass0 = list(
    poly_degree = 2
  ),
  
  opts_hrf = list(
    hrf_fir_taps = 12,
    good_voxel_R2_threshold = 0.05
  ),
  
  # ... and so on for other modules
)

# Use in NDX_Process_Subject
results <- NDX_Process_Subject(
  Y_fmri = fmri_data,
  events = event_data,
  motion_params = motion_data,
  run_idx = run_indices,
  TR = 2.0,
  user_options = user_options
)
```

## Module-Specific Options

### Initial GLM (Pass 0)

The `opts_pass0` controls the initial GLM for residual generation:

```{r eval=FALSE}
opts_pass0 = list(
  poly_degree = 2  # Controls polynomial drift regressors
)
```

### HRF Estimation

The `opts_hrf` list controls the HRF estimation process, including FIR parameters, clustering options, and constraints:

```{r eval=FALSE}
opts_hrf = list(
  hrf_fir_taps = 12,               # Number of FIR taps
  hrf_fir_span_seconds = 24,       # Span of FIR in seconds
  good_voxel_R2_threshold = 0.05,  # R² threshold for good voxels
  cv_folds = 5,                    # Cross-validation folds
  hrf_cluster_method = "none",     # Clustering method (none, kmeans, pam)
  num_hrf_clusters = 1,            # Number of HRF clusters
  hrf_cone_nonneg = TRUE,          # Non-negativity constraint
  hrf_cone_unimodal = TRUE,        # Unimodality constraint
  hrf_cone_normalize_area = TRUE   # Normalize area under curve
)
```

### RPCA Component Extraction

The `opts_rpca` list controls Robust PCA for temporal noise component extraction:

```{r eval=FALSE}
opts_rpca = list(
  k_global_target = 20,           # Target global components 
  k_elbow_drop_ratio = 0.02,      # Elbow detection parameter
  k_rpca_min = 10,                # Minimum components
  k_rpca_max = 50                 # Maximum components
)
```

### Spectral Component Extraction

The `opts_spectral` list controls spectral analysis for frequency-based noise components:

```{r eval=FALSE}
opts_spectral = list(
  n_sine_candidates = 10,         # Number of sine candidates
  nyquist_guard_factor = 0.9,     # Nyquist guard factor 
  k_tapers = 5,                   # Number of tapers
  nw = 3                          # Time-bandwidth product
)
```

### AR Whitening

The `opts_whitening` list controls the pre-whitening process:

```{r eval=FALSE}
opts_whitening = list(
  order = 2,                      # AR order
  global_ar_on_design = TRUE      # Apply global AR on design
)
```

### Ridge Regression

The `opts_ridge` list controls the ridge regression for final denoising:

```{r eval=FALSE}
opts_ridge = list(
  anisotropic_ridge_enable = TRUE,  # Enable anisotropic ridge
  lambda_perp_signal = 0.1,         # Lambda for signal components
  lambda_noise_rpca = 1.0,          # Lambda for RPCA noise
  lambda_noise_spectral = 1.0       # Lambda for spectral noise
)
```

### Annihilation Mode

The `opts_annihilation` list controls the Annihilation Mode, which integrates GLMdenoise-like functionality:

```{r eval=FALSE}
opts_annihilation = list(
  annihilation_enable_mode = TRUE,               # Enable Annihilation Mode
  annihilation_gdlite_k_max = 30,                # Maximum number of GLMdenoise PCs
  annihilation_gdlite_r2_thresh_noise_pool = 0.05  # R² threshold for noise pool
)
```

## Advanced Usage Examples

### Conservative Denoising

```{r eval=FALSE}
conservative_options <- list(
  max_passes = 2,
  opts_rpca = list(k_global_target = 10, k_rpca_max = 20),
  opts_ridge = list(lambda_perp_signal = 0.05, lambda_noise_rpca = 0.5)
)
```

### Aggressive Denoising

```{r eval=FALSE}
aggressive_options <- list(
  max_passes = 5,
  min_des_gain_convergence = 0.001,
  opts_rpca = list(k_global_target = 30, k_rpca_max = 100),
  opts_ridge = list(lambda_perp_signal = 0.2, lambda_noise_rpca = 5.0)
)
```

### Annihilation Mode Configuration

```{r eval=FALSE}
annihilation_options <- list(
  opts_annihilation = list(
    annihilation_enable_mode = TRUE,
    annihilation_gdlite_k_max = 25,
    annihilation_gdlite_r2_thresh_noise_pool = 0.05,
    annihilation_gdlite_tsnr_thresh_noise_pool = 30
  ),
  
  # Special ridge settings for Annihilation Mode
  opts_ridge = list(
    lambda_noise_gdlite = 1.0,      # Lambda for GLMdenoise PCs
    lambda_noise_ndx_unique = 0.5   # Lambda for unique ND-X components
  )
)
```

## Troubleshooting

### Common Issues

- **Error: Parameter X is not valid**: Check the parameter name and make sure it's nested in the correct sub-list.
- **Warning: Using default for Y**: The option wasn't found and a default was used. Check your option structure.

### Finding Default Values

All default values are documented in the `?ndx_user_options` help page. You can also inspect the code for specific functions to see how defaults are implemented.

## Best Practices

1. **Start with Defaults**: Run with minimal options first, then customize based on your needs.
2. **Monitor Diagnostics**: Use the HTML diagnostics to see how changes affect performance.
3. **Incremental Changes**: Change one set of options at a time to understand effects.
4. **Document Your Settings**: Save your `user_options` along with your results for reproducibility.

## References

For additional information, refer to:

- The function documentation (`?NDX_Process_Subject`)
- The options reference (`?ndx_user_options`)
- The ND-X methodology paper (citation to be added) 